
\chapter{Acknowledgements}

I have been working on this project since I was a full-time research assistant
at Formal Methods Laboratory in Academia Sinica.
I should like to thank Prof. Yu-Fang Chen and Prof. Bow-Yaw Wang.
The whole framework is gradually developed under their guidance.
Their solid training and teaching also helped me establish my knowledge of
Software Verification and Formal Methods.
In particular, the opportunity and the experience of attending international
conferences surely is very enlightening though quite frightening.
In addition, I am specially grateful to and surprised by the assistance from
Dr. Ming-Hsien Tsai on the implementation.
His efficiency and proficiency on software programming skills are simply
remarkable and incomparable.
Without his assistance, I am sure I cannot finish the implementation on my own,
let alone participating the competition.

I also learned a lot and pretty enjoyed my life in Software Testing Laboratory.
Prof. Farn Wang is a great role model as a academic researcher.
His ambition on establishing an industrial strength software testing tool is
respectable.
The laboratory members and assistants are just adorable.
I need not mention your achievements on researches.
You guys inspired me in a unique way.
Even when we were both struggling for homeworks, reports, exams, and finally
the graduation, you guys still found many ways to relax and relief the tension
among the whole lab.
"Work Hard, Play Hard" is the spirit I learned.
You guys rock.

Finally, my appreciation to my parents and siblings is beyond words.
Thanks for your constant nagging on me about my health, both physical and
mental, as well as trying to stop my 26-year-without-girlfriend achievement.
Your concerns and considerations let me understand that,
even if I might fail so hard and be so frustrated,
I can go home for recovery and face the challenge next time.

\vspace{8mm}

\noindent
Thank you all.

\vspace{8mm}

\noindent
National Taiwan University, Taipei

\noindent
July, 2015 \hfill 謝橋 \ Hsieh, Chiao

