%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classicthesis-Styled CV
% LaTeX Template
% Version 1.0 (22/2/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro Plasmati
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%    PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[letterpaper, UTF8]{scrartcl}

\usepackage{comment}

\reversemarginpar % Move the margin to the left of the page



\newcommand{\MarginText}[1]{\marginpar{\raggedleft\itshape\small#1}} % New command defining the margin text style

\usepackage[nochapters]{classicthesis} % Use the classicthesis style for the style of the document

\setlength{\marginparsep}{5mm}

\usepackage[letterpaper, top=20mm, bottom=20mm, outer=18mm,inner=33mm]{geometry}
\usepackage[LabelsAligned]{currvita} % Use the currvita style for the layout of the document

\renewcommand{\cvheadingfont}{\LARGE\color{Maroon}} % Font color of your name at the top

\usepackage{doi}
\usepackage{hyperref} % Required for adding links    and customizing them
\hypersetup{colorlinks, breaklinks, urlcolor=Maroon, linkcolor=Maroon} % Set link colors

\date{}

\newlength{\datebox}\settowidth{\datebox}{Spring 2011} % Set the width of the date box in each block

\newcommand{\NewEntry}[3]{\noindent\hangindent=0em\hangafter=0 \parbox{\datebox}{\small \textit{#1}} {\small #2} #3 % Define a command for each new block - change spacing and font sizes here: #1 is the left margin, #2 is the italic date field and #3 is the position/employer/location field
\vspace{0.3em}} % Add some white space after each new entry

\newcommand{\Description}[1]{\hangindent=0em\hangafter=0\noindent\raggedright\footnotesize{#1}\par\normalsize\vspace{0.3em}}
 % Define a command for descriptions of each entry - change spacing and font
 % sizes here
 
\newcommand{\newsectsep}{\vspace{0.9em}}
\newcommand{\reducesep}{\vspace{-0.3em}}

\newcommand{\papertitle}[1]{{``#1''}}

\newcounter{papers}

%----------------------------------------------------------------------------------------
\usepackage{ctex}

\begin{document}

\thispagestyle{empty} % Stop the page count at the bottom of the first page

%----------------------------------------------------------------------------------------
%    NAME AND CONTACT INFORMATION SECTION
%----------------------------------------------------------------------------------------
\begin{CJK}{UTF8}{kai}

\begin{cv}{謝~橋/Hsieh, Chiao/シェ~チャオ}\vspace{1em} % Your name

% Email address and Personal website
\Description{%
Email: \href{mailto:hsieh.chiao.7k@kyoto-u.ac.jp}{hsieh.chiao.7k@kyoto-u.ac.jp} $\cdot$
LinkedIn: \url{https://linkedin.com/in/hc825b} $\cdot$
Website: \url{https://hc825b.github.io}
}

%\NewEntry{phone}{(H) +1 (000) 111 1111\ \ $\cdotp$\ \ (M) +1 (000) 111 1112} % Phone number(s)

\newsectsep % Extra white space between the personal information section and goal

\noindent\spacedlowsmallcaps{Research Interest}\vspace{0.5em} % Goal heading, could be used for a quotation or short profile instead

\Description{Software Verification and Testing $\cdot$
    Formal Methods $\cdot$
    Cyber-physical System Verification
}% Goal text

\Description{Topics: Assuring and certifying safety of cyber-physical systems containing complex or incompletely specified components.
My main insight is to develop search algorithms for approximate abstractions of these components so that the approximations are tractable for existing formal verification techniques.
In particular, I have been studying systems with vision-based perception, blackbox dynamics, and distributed coordination.
}\vspace{1em} % Goal text

%----------------------------------------------------------------------------------------
%    GRANTS
%----------------------------------------------------------------------------------------

\spacedlowsmallcaps{Ongoing Research Projects}\vspace{0.5em}

\NewEntry{FY 2024 -- 2025}{\MarginText{科研費\\KAKENHI}\papertitle{Safety Verification of Black Box Cyber-Physical Systems via Lyapunov Stability Certificates}

\Description{%
	Category: JSPS Grant-in-Aid for Research Activity Start-Up (Grant \# \href{https://kaken.nii.ac.jp/en/grant/KAKENHI-PROJECT-24K23861}{24K23861})}
}

%----------------------------------------------------------------------------------------
%    EDUCATION
%----------------------------------------------------------------------------------------

\spacedlowsmallcaps{Education}\vspace{0.5em}

\NewEntry{2015 -- 2023}{\MarginText{CS@Illinois} Doctor of Philosophy}

\Description{\textbf{Reliable Autonomy Group} in Department of Computer Science at University of Illinois Urbana-Champaign\newline
    Thesis: \papertitle{Abstractions for Safety Assurance of Autonomous Systems}\newline
    Advisor:
    Prof. Mitra, Sayan $\cdot$ \href{mailto:mitras@illinois.edu}{mitras@illinois.edu}%
}

%------------------------------------------------

\NewEntry{2013 -- 2015}{\MarginText{臺大電子所}Master of Science}

\Description{\textbf{Verification Automation \& Software Testing Lab}
    in Graduate Institute of Electronics Engineering at National Taiwan University\newline
    Thesis: \papertitle{Verifying Recursive Programs Using Intraprocedural Analyzers}\newline
    Advisors:
    Prof. Chen, Yu-Fang $\cdot$ \href{mailto:yfc@iis.sinica.edu.tw}{yfc@iis.sinica.edu.tw},
    Prof. Wang, Farn $\cdot$ \href{mailto:farn@ntu.edu.tw}{farn@ntu.edu.tw}
}

%------------------------------------------------

\NewEntry{2007 -- 2011}{\MarginText{臺大電機系}Bachelor of Science}

\Description{\textbf{Design Verification Lab}
    in Department of Electrical Engineering at National Taiwan University\newline
    Projects: \papertitle{Symbolic model checking on SystemC designs} with
    Prof. Huang, Chung-Yang $\cdot$ \href{mailto:cyhuang@ntu.edu.tw}{cyhuang@ntu.edu.tw}}

%------------------------------------------------

\newsectsep % Extra space between major sections

%----------------------------------------------------------------------------------------
%    WORK EXPERIENCES
%----------------------------------------------------------------------------------------

\spacedlowsmallcaps{Work Experiences}\vspace{0.5em}

\NewEntry{2023 -- Today}{\MarginText{京都大学\\情報学研究科} Postdoc in \textbf{Computer Software Group} at Graduate School of Informatics, Kyoto University}

\Description{%\textbf{Computer Software Group},
%    Prof. Suenaga, Kohei $\cdot$ \href{mailto:ksuenaga@fos.kuis.kyoto-u.ac.jp}{ksuenaga@fos.kuis.kyoto-u.ac.jp}
}

\NewEntry{2018 -- 2023}{\MarginText{CSL@Illinois} Graduate Research Assistant in \textbf{Reliable Autonomy Group} at Coordinated Science Laboratory, UIUC}

%\Description{%\textbf{Reliable Autonomy Group},
%	Prof. Mitra, Sayan $\cdot$ \href{mailto:mitras@illinois.edu}{mitras@illinois.edu}
%}

\NewEntry{2017 -- 2018}{\MarginText{CS@Illinois} TA at Department of Computer Science, UIUC}

\Description{%\textbf{CS 527}
%    Prof. Marinov, Darko $\cdot$ \href{mailto:marinov@illinois.edu}{marinov@illinois.edu},
%    \textbf{CS 421} Prof. Beckman, Mattox $\cdot$ \href{mailto:mattox@illinois.edu}{mattox@illinois.edu}
}

\NewEntry{Summer 2016}{\MarginText{IBM Research}Research Intern in \textbf{Cloud DevOps and Cognitive Platform group} at IBM Research, New York}

\Description{%\textbf{Cloud DevOps and Cognitive Platform group},
%	Dr. Nagpurkar, Priya $\cdot$
%	\href{mailto:pnagpurkar@us.ibm.com}{pnagpurkar@us.ibm.com}
}


\NewEntry{2015 -- 2017}{\MarginText{CS@Illinois} RA in \textbf{Automated Software Engineering Group} at Department of Computer Science, UIUC}

\Description{%\textbf{Automated Software Engineering Group},
%	Prof. Xie, Tao $\cdot$ \href{mailto:taoxie@pku.edu.cn}{taoxie@pku.edu.cn}
}

%------------------------------------------------

\NewEntry{2012 -- 2015}{\MarginText{中研院資訊所} RA in \textbf{Programming Languages and Formal Methods Group} at Institute of Information Science, Academia Sinica, Taipei}

\Description{%\textbf{Programming Languages and Formal Methods Group},
%	Prof. Chen, Yu-Fang $\cdot$ \href{mailto:yfc@iis.sinica.edu.tw}{yfc@iis.sinica.edu.tw}
}

%------------------------------------------------

\NewEntry{2011 -- 2012}{\MarginText{海軍}Secondary Lieutenant. Compulsory Military Service at Republic of China Navy}
\newsectsep % Extra space between major sections

%----------------------------------------------------------------------------------------
%    SERVICES
%----------------------------------------------------------------------------------------

\spacedlowsmallcaps{Services and Volunteering}\vspace{0.5em}

\Description{2024 $\cdot$ Fast Abstracts Track Committee at 35\textsuperscript{th} Intl. Symp. Software Reliability Engineering~(ISSRE~2024)}

\reducesep

\Description{2022 $\cdot$ Artifact Evaluation Committee at 2022 European Conf. Object-Oriented Programming~(ECOOP~2022)}

\reducesep

\Description{2022 $\cdot$ Student Volunteer at 34\textsuperscript{th} Intl. Conf. Computer Aided Verification~(CAV~2022)}

\reducesep

\Description{2018 $\cdot$ Graduate Mentor in Promoting Undergraduate Research in Engineering~(PURE) at UIUC}

\reducesep

\Description{2017 $\cdot$ Student Volunteer at 32\textsuperscript{nd} Intl. Conf. Automated Software Engineering~(ASE~2017)}


%----------------------------------------------------------------------------------------
%    PUBLICATIONS
%----------------------------------------------------------------------------------------

\spacedlowsmallcaps{Journal Articles}\vspace{0.5em}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
    \papertitle{Perception Contracts for Safety of ML-Enabled Systems}\newline
    Angello~Astorga*, \underline{Chiao~Hsieh}*, P.~Madhusudan, Sayan~Mitra.
    \textit{Proc. ACM on Programming Languages (PACMPL), vol. 7, iss. OOPSLA2, Article No. 232, Oct. 2023}. \doi{10.1145/3622875}\newline
    *Equal contributions as co-first authors.\newline
    \textreferencemark This paper is accepted by OOPSLA, a CORE rank ``A" conference.
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
    \papertitle{Verifying Controllers with Vision-based Perception using Safe Approximate Abstractions}\newline
    \underline{Chiao~Hsieh}, Yangge~Li, Dawei~Sun, Keyur~Joshi, Sasa~Misailovic, Sayan~Mitra.
    \textit{IEEE Trans. Computer-Aided Design of Integrated Circuits and Systems (TCAD), vol. 41, no. 11, Nov. 2022}. \doi{10.1109/TCAD.2022.3197508}\newline
    \textreferencemark This paper is accepted by EMSOFT, a CORE rank ``A" conference.
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
    \papertitle{Koord: A Language for Programming and Verifying Distributed Robotics Application.}\newline
    Ritwika~Ghosh, \underline{Chiao~Hsieh}, Sasa~Misailovic, Sayan~Mitra.
    \textit{Proc. ACM on Programming Languages (PACMPL), vol. 4, iss. OOPSLA, Article No. 232, Nov. 2020}.
    \doi{10.1145/3428300}\newline
    \textreferencemark This paper is accepted by OOPSLA, a CORE rank ``A" conference.
}


\spacedlowsmallcaps{Conference Publications}\vspace{0.5em}


\stepcounter{papers}
\Description{\MarginText{\thepapers}%
	\papertitle{GAS: Generating Fast and Accurate Surrogate Models for Autonomous Vehicle Systems}\newline
	Keyur~Joshi, \underline{Chiao~Hsieh}, Sayan~Mitra, Sasa~Misailovic.
	\textit{To appear in 35\textsuperscript{th} Intl. Symp. Software Reliability Engineering (ISSRE~2024)}. preprint \doi{10.48550/arXiv.2208.02232}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
	\papertitle{Assuring Safety of Vision-based Swarm Formation Control}\newline
	\underline{Chiao~Hsieh}, Yangge~Li, Yubin Koh, Sayan~Mitra.
	\textit{Proc. 2024 American Control Conf. (ACC~2024)}. \doi{10.23919/ACC60939.2024.10644491}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
    \papertitle{Industry-track: Challenges in Rebooting Autonomy with Deep Learned Perception}\newline
    Michael Abraham, Aaron Mayne, Tristan Perez, Italo Romani De Oliveira, Huafeng Yu,
    \underline{Chiao~Hsieh}, Yangge~Li, Dawei~Sun, and Sayan~Mitra.
    \textit{Proc. 2022 Intl. Conf. Embedded Software (EMSOFT~2022)}.
    \doi{10.1109/EMSOFT55006.2022.00016}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
    \papertitle{Programming Abstractions for Simulation and Testing on Smart Manufacturing Systems}\newline
    \underline{Chiao~Hsieh}, Daniel Wu, Yubin Koh, Sayan~Mitra.
    \textit{Proc. 18\textsuperscript{th} Intl. Conf. Automation Science and Engineering (CASE~2022)}.
    \doi{10.1109/CASE49997.2022.9926564}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{SkyTrakx: A Toolkit for Simulation and Verification of Unmanned Air-Traffic Management Systems.}\newline
\underline{Chiao~Hsieh}, Hussein~Sibai, Hebron~Taylor, Yifeng~Ni, Sayan~Mitra.
\textit{Proc. 2021 IEEE Intl. Intelligent Transportation Systems Conf. (ITSC~2021)}.
\doi{10.1109/ITSC48978.2021.9564492}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{CyPhyHouse: A Programming, Simulation, and Deployment Toolchain for Heterogeneous Distributed Coordination.}\newline
Ritwika Ghosh, Joao~P. Jansch-Porto, \underline{Chiao Hsieh}, Amelia Gosse, Minghao Jiang, Hebron Taylor, Peter Du, Sayan Mitra, Geir Dullerud.
\textit{Proc. 2020 IEEE Intl. Conf. Robotics and Automation (ICRA~2020)},
\doi{10.1109/ICRA40945.2020.9196513}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{Dione: A Protocol Verification System Built with Dafny for I/O Automata.}\newline
\underline{Chiao~Hsieh}, Sayan Mitra.
\textit{Proc. 15\textsuperscript{th} Intl. Conf. integrated Formal Methods (iFM~2019)}.
\doi{10.1007/978-3-030-34968-4\_13}
}

%------------------------------------------------

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{PAC Learning-Based Verification and Model Synthesis.}\newline
Yu-Fang~Chen, \underline{Chiao~Hsieh}, Ond\v{r}ej~Leng\'{a}l, Tsung-Ju~Lii, Ming-Hsien~Tsai, Bow-Yaw~Wang, Farn~Wang.
\textit{Proc. 38\textsuperscript{th} Intl. Conf. Software Engineering (ICSE~2016)}.
\doi{10.1145/2884781.2884860}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{CPArec: Verifying Recursive Programs via Source-to-Source Program Transformation - (Competition Contribution).}\newline
Yu-Fang~Chen, \underline{Chiao~Hsieh}, Ming-Hsien~Tsai, Bow-Yaw~Wang, Farn~Wang.
\textit{Proc. 21\textsuperscript{st} Intl. Conf. Tools and Algorithms for the Construction and Analysis of Systems (TACAS~2015)}.
\doi{10.1007/978-3-662-46681-0\_35}
}

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{Verifying Recursive Programs Using Intraprocedural Analyzers.}\newline
Yu-Fang~Chen, \underline{Chiao~Hsieh}, Ming-Hsien~Tsai, Bow-Yaw~Wang, Farn~Wang.
\textit{Proc. 21\textsuperscript{st} Intl. Static Analysis Symposium (SAS~2014)}.
\doi{10.1007/978-3-319-10936-7\_8}
}

%------------------------------------------------

\stepcounter{papers}
\Description{\MarginText{\thepapers}%
\papertitle{Symbolic model checking on SystemC designs.}\newline
Chun-Nan~Chou, Yen-Sheng~Ho, \underline{Chiao~Hsieh}, Chung-Yang (Ric) Huang.
\textit{Proc. 49\textsuperscript{th} Annual Design Automation Conf. (DAC 2012)}.
\doi{10.1145/2228360.2228421}
}

%------------------------------------------------

% \newsectsep % Extra space between major sections

%----------------------------------------------------------------------------------------
%    COMPUTER SKILLS
%----------------------------------------------------------------------------------------

%\spacedlowsmallcaps{software/hardware development skills}\vspace{0.5em}

% \Description{\MarginText{Familiar With}C/C++ $\cdot$ Python $\cdot$ Git $\cdot$
%	Bash $\cdot$ Ubuntu Linux $\cdot$ \LaTeX $\cdot$ Dafny $\cdot$ Robot Operating System $\cdot$ Gazebo simulator $\cdot$
%	Subversion $\cdot$ AirSim simulator}

% \Description{\MarginText{Basic}Haskell $\cdot$ K Semantic Framework}

%------------------------------------------------

% \newsectsep % Extra space between major sections

%----------------------------------------------------------------------------------------
%    TEACHING EXPERIENCE
%----------------------------------------------------------------------------------------

\noindent\spacedlowsmallcaps{Teaching Experience}\vspace{0.5em}

\NewEntry{Spring 2018}{\MarginText{CS@Illinois}TA for CS 421 Programming Languages and Compilers}
\reducesep

%\Description{%
%    Prof. Beckman, Mattox $\cdot$
%    \href{mailto:mattox@illinois.edu}{mattox@illinois.edu}}

\NewEntry{Fall 2017}{TA for CS 527 Topics in Software Engineering}
\reducesep

%\Description{%
%    Prof. Marinov, Darko $\cdot$
%    \href{mailto:marinov@illinois.edu}{marinov@illinois.edu}}


\NewEntry{Summer 2015}{\MarginText{臺大} TA for Formosan Summer School on Logic, Language, and Computation~(FLOLAC~2015)}
\reducesep

%\Description{%
%    Prof. Chen, Yu-Fang $\cdot$
%    \href{mailto:yfc@iis.sinica.edu.tw}{yfc@iis.sinica.edu.tw}}

\newsectsep % Extra space between major sections


%------------------------------------------------

% \newsectsep % Extra space between major sections

%----------------------------------------------------------------------------------------
%    OTHER INFORMATION
%----------------------------------------------------------------------------------------
\begin{comment}
\spacedlowsmallcaps{Other Information}\vspace{0.5em}

%------------------------------------------------
\Description{\MarginText{Presentations}2021 $\cdot$ 15 min. presentation for SkyTrakx at ITSC~2021}

\reducesep

\Description{2020, 2021 $\cdot$ 20 min. presentation for Koord at OOPSLA~2020 and SPLASH~2021}

\reducesep

\Description{2019 $\cdot$ 30 min. presentation for Dione at iFM~2019}

\reducesep

\Description{2015 $\cdot$ 5 min. tool paper presentation at TACAS~2015}

\reducesep

\Description{2014 $\cdot$ 20 min. research paper presentation at SAS 2014}



%------------------------------------------------

\newlength{\langbox} % Create a new length for the length of languages to keep them equally spaced
\settowidth{\langbox}{English} % Length equals the length of "English" - if you have a longer language in your list put it here

\Description{\MarginText{Languages}\parbox{\langbox}{\textsc{Chinese}} $\cdot$ Mother tongue}

\reducesep

\Description{\parbox{\langbox}{\textsc{English}} $\cdot$ Professionally Fluent}

\reducesep

\Description{\parbox{\langbox}{\textsc{Japanese}} $\cdot$ Basic (can read news passages)}
\end{comment}
%------------------------------------------------

\end{cv}
\end{CJK}

\end{document}
